\chapter{Tests}

\section{Overview}

To better investigate the impact on the community, a questionnaire has been sent and filled by various active members of the SourceMod community.
It was distributed on the official AlliedModders forums, promoted on the SourceMod Discord server, and sent to revelant people willing to provide their answers.
Those answers were collected from 44 of them, where 5 were server owners, 6 were developers and the remaining 33 were both an owner and a developer.
Subjects varied in their involvement and time in the community, some being members less than a year, while others where for over 9 years.

A special thanks goes to two moderators of the Discord server and heads of the community: \textbf{asherkin}, also the owner of LimeTech, and \textbf{headline}.
They helped in getting this questionnaire to more people than a single person could.

\section{Questions}

\begin{enumerate}
    \item \textbf{Are you a developer and/or a server owner?}

    SourceMod Addon Manager is a tool aimed at helping those two kinds of users.
    The end users -- players -- are not affected by it.
    It is worth distinguishing which of them is answering the questionnaire.

    \item \textbf{How long have you been a part of the SourceMod community?}

    The community has its years and it is important to roughly establish when the subject had joined.
    Many of the tools listed in the introduction were not available back in the day, and different users are used to different solutions.

    \item \textbf{Are you familiar with package managers and ever used one?}

    An understandably popular platform for servers is Linux, where package managers are widespread.
    Due to this, most developers and server owners are accustomed with the concept of automatically installing software packages.
    Yet, a small amount of community might be unaware of this, as game server rentals abstract the platform and file management entirely.

    \item \textbf{Do you own a SourceMod server? (ignore if not)}

    This multi-choice question overviews different means of owning a server.
    The subject may answer that they either own a dedicated machine, rent a game server only, or self host it.
    The latter is popular among developers, while the former is more popular among server owners instead.
    Lastly, renting a game server is cheaper, but much less scalable in the long run.
    It's this group that SourceMod Addon Manager cannot help yet.

    \item \textbf{Do you feel the SourceMod community needs a package manager?}

    The community is used to manually install addons by their own means.
    Answers to this question range from simply ``no,'' to ``yes, very much.''

    \item \textbf{Would you use a package manager?}

    To complement the previous question, this one relates to the user themselves.
    Likewise, answers range from ``no,'' to ``yes, very often.''

    \item \textbf{What do you find most problematic when you want to install a plugin?}

    Subjects were able to provide their own answer to this question or skip it altogether.

    \item \textbf{What is your most important feature in a fully functioning package manager for SourceMod, besides automated installation and removal?}

    Subjects were selecting from a predetermined list of answers or were able to provide their own.
    It is plausible that the SourceMod community is looking for something more specific than a standard package manager.
\end{enumerate}

\section{Results}

Below are the questions themselves and their compiled results.

\subsubsection{Are you a developer and/or a server owner?}

\subsubsection{How long have you been a part of the SourceMod community?}

\subsubsection{Are you familiar with package managers and ever used one?}

\subsubsection{Do you own a SourceMod server? (ignore if not)}

\subsubsection{Do you feel the SourceMod community needs a package manager?}

\subsubsection{Would you use a package manager?}

\subsubsection{What do you find most problematic when you want to install a plugin?}
